### 2025-05-01 docs/Homeownership_BankOfMomAndDad.tex
Updated empirics, moved rent-to-own reg to appendix

I now utilize microdata from the PSID to examine the empirical association between transfer receipt and parental financial resources on housing outcomes. The sample is limited to households aged 25 to 44 to focus on the housing choices of young adults. I use the sum of parental net worth and income (denoted as wealth) to measure parents' financial resources in all specifications for two reasons. First, this is consistent with models of altruism, as the one used later in the paper. Second, since parents' wealth and income are highly correlated and I have few observations---fewer than 1,000 for many specifications---the standard errors are large when both are included.

\subsubsection{Transfers and the Rent-to-Own Transition}

While a growing literature documents that parental transfers increase the probability that renters become homeowners (e.g.\ \citealp{wold2024housing}; \citealp{Blickle2019}; \citealp{benetton2022dynastic}), I replicate this result to highlight the importance of transfers for housing outcomes. I estimate
\begin{equation}	\label{eq:transferrent}
	\Pr\left(Own_{i,t+2}=1 \mid Own_{i,t}=0\right)
	= \alpha + \beta\,Transfer_{i,t+2}
	  + \gamma' X_{i,t} + \varepsilon_{i,t},
\end{equation}
where \(Own_{i,t}=0\) denotes a renter at time \(t\), \(Transfer_{i,t}=1\) if the household received a transfer between survey waves, and \(X_{i,t}\) includes log parental wealth, log household income and log household net worth, log household size, age, parental age, and dummies for  education, marital status, race, household composition changes, and state and year fixed effects. The sample is limited to renting households aged 25 to 44 who have never been observed as homeowners. Table \ref{tab:newowners_short} reports the estimated coefficient on transfer ($\beta$), with the full coefficient table relegated to Appendix Table \ref{tab:newowners}. These regressions are similar to those in \cite{Lee2018}, who also use the PSID.  It is important to note that transfer receipt is not exogenous---many transfers may be made precisely because the renter is buying a home.

\begin{table}[tb]
	\centering
	\begin{threeparttable}
		\caption{The Transition to Ownership}
		\label{tab:newowners_short}
		\small 
				\input{../tabfig/regr/newowners_short.tex}

	\end{threeparttable}
	{\begin{footnotesize}\begin{flushleft}\vspace{-0.1in}%
		\textit{Notes:} Standard errors in parentheses, clusted at the household level. \textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
	\end{flushleft}\end{footnotesize}}		
\end{table}

@@

@@

To understand which variables predicts retainment of ownership, I estimate
\begin{equation}	\label{eq:regmaintain}
	\Pr\left(Own_{i,t+\Delta}=1|Own_{i,t}=1\right)
	= \beta\ln(NetWorth_{i,t}) + \zeta\ln(ParWealth_{i,t})
	  + \gamma' X_{i,t} + \varepsilon_{i,t},
\end{equation}
where $Own_{i,t+\Delta}=1$ denotes an owner at time $t+\Delta$, $\beta$ and $\zeta$ measures the effect of log household net worth and parental wealth, respectively  and $X_{i,t}$ is a vector of control variables. The only changes relative to Eq. \ref{eq:transferrent} is that the outcome variable conditions on current ownership and measures future ownership, that there are no controls for transfer receipt, and that I have taken the two variables of interest out of the vector of control variables. The sample is limited to households aged 25 to 44. These regressions are motivated by \cite{bond2021role}, who estimate similar equations in the PSID and the Health and Retirement Study. Figures \ref{fig:coef_allowner} and \ref{fig:coef_ftowner} plots the two coefficients of interest over different time horizons.

@@

To examine what predicts mortgage delinquency, I estimate: 
\begin{equation*}
\Pr(Behind_{i,t+2}) = \beta\ln(NetWorth_{i,t}) + \zeta\ln(ParWealth_{i,t}) + \gamma' X_{i,t} + \varepsilon_{i,t},
\end{equation*}
using the same control variables as in the previous regression. Table~\ref{tab:hypo} reports the estimated coefficients on household net worth ($\beta$); full results appear in Appendix Table~\ref{tab:hypo_long}.

@@

\begin{table}
	\centering
	\begin{threeparttable}
		\caption{Housing Choices and Parental Wealth}
		\label{tab:hypo}
		\small 
				\input{../tabfig/regr/hypoII_short.tex}

	\end{threeparttable}
	{\begin{footnotesize}\begin{flushleft}
		\textit{Notes:} Standard errors in parentheses, clustered at the household level. \textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
		\end{flushleft}\end{footnotesize}}	
\end{table}

@@

\subsubsection{Households with Wealthier Parents Are Less Likely to Downsize during Unemployment Spells}\label{sec:eventstudy}
Distressed and forced home sales are costly and often triggered by unemployment \citep{kermani2021racial,hsu2018unemployment}. To test whether parental wealth mitigates this risk, I perform a simple event study of how unemployment affects the likelihood of downsizing.

@@

\subsection{Taking Stock of the Data}
Taking stock, we have seen that parental transfers have become more important for young homeowners over the past two decades. About 30\% of young households receive transfers from their parents in a given year, and recipients' parents are wealthier. Recipients are also more likely to transition from renting to owning, also when we condition on a rich set of covariates. Parental wealth is also positively associated with maintaing homeownership. The probability of mainting homeownership is increasing in parental wealth, households with richer parents are less likely to behind on mortgages and less likely to downsize during unemployment sptells. These results suggest that homeowners with wealthy parents find illiquidity of homeownership less problematic.

@@

>>>>

@@

I now utilize microdata from the PSID to examine the empirical association between parental financial resources on housing outcomes. The sample is limited to households aged 25 to 44 to focus on the housing choices of young adults. I use the sum of parental net worth and income (denoted as wealth) to measure parents' financial resources in all specifications for two reasons. First, this is consistent with models of altruism, as the one used later in the paper. Second, since parents' wealth and income are highly correlated and I have few observations---fewer than 1,000 for many specifications---the standard errors are large when both are included.

\subsubsection{Transfers and the Transition into Homeownership}
Since many papers have already estimated the relationship between parental transfers and transitions into homeownership (e.g.\ \citealp{wold2024housing}; \citealp{Blickle2019}; \citealp{benetton2022dynastic}; \citealp{Lee2018}), I replicate these regressions to highlight the association between transfer receipt and rent-to-own transitions, but relegate the full results to Appendix~\ref{app:rent_to_own}. Receiving a parental transfer above \$500 is associated with a 1.2 percentage point increase in the probability of becoming a homeowner (not statistically significant), after controlling for household characteristics. Receiving large transfers above \$10,000 is associated with a 7.9 percentage point increase in the probability of becoming a homeowner ($p<0.1$), even after including household fixed effects. These are sizable effects relative to the baseline rent-to-own transition probability of about 18\% during the sample period.

@@

To understand which variables predicts retainment of ownership, I estimate
\begin{equation}	\label{eq:regmaintain}
	\Pr\left(Own_{i,t+\Delta}=1|Own_{i,t}=1\right)
	= \beta\ln(NetWorth_{i,t}) + \zeta\ln(ParWealth_{i,t})
	  + \gamma' X_{i,t} + \varepsilon_{i,t},
\end{equation}
where $Own_{i,t+\Delta}=1$ denote ownership at time $t+\Delta$, $\beta$ and $\zeta$ measure the effects of log household net worth and parental wealth, respectively. The vector of control variables  $X_{i,t}$ consists of log household income, log household size, age, parental age, and dummies for  education, marital status, race, household composition changes, and state and year fixed effects.  The sample is limited to households aged 25 to 44. These regressions are motivated by \cite{bond2021role}, who estimate similar equations in the PSID and the Health and Retirement Study. Figures \ref{fig:coef_allowner} and \ref{fig:coef_ftowner} plot the two coefficients of interest over different time horizons.

@@

To examine what predicts mortgage delinquency, I estimate: 
\begin{equation}\label{eq:regbehind}
\Pr(Behind_{i,t+2}) = \beta\ln(NetWorth_{i,t}) + \zeta\ln(ParWealth_{i,t}) + \gamma' X_{i,t} + \varepsilon_{i,t},
\end{equation}
using the same control variables as in the previous regression. Table~\ref{tab:hypo} reports the estimated coefficients on parental wealth ($\zeta$); full results appear in Appendix Table~\ref{tab:hypo_long}.

@@

\begin{table}
	\centering
	\begin{threeparttable}
		\caption{Parental Wealth and Future Mortgage Delinquencies}
		\label{tab:hypo}
		\small 
				\input{../tabfig/regr/hypoII_short.tex}

	\end{threeparttable}
	{\begin{footnotesize}\begin{flushleft}
		\textit{Notes:} The estimated association between parental wealth and the probability of behind on mortgage payments in the next period from regressions using specification \ref{eq:regbehind}. See Appendix Table~\ref{tab:hypo_long} for all results. Standard errors in parentheses, clustered at the household level. \textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
		\end{flushleft}\end{footnotesize}}	
\end{table}

@@

\subsubsection{Owners with Wealthy Parents Don't Downsize in Unemployment}\label{sec:eventstudy}
Distressed and forced home sales are costly and often triggered by unemployment \citep{kermani2021racial,hsu2018unemployment}. To test whether parental wealth mitigates this risk, I perform a simple event study of how unemployment affects the likelihood of downsizing.

@@

\subsection{Taking Stock of the Data}
Taking stock, we have seen that parental transfers have become more important for young homeowners over the past two decades. About 20\% of young households receive transfers from their parents in a given year, and recipients have wealthier parents. Recipients are more likely to transition from renting to owning, even after conditioning on a rich set of covariates. Parental wealth is also positively associated with maintaining homeownership. Households with richer parents are less likely to fall behind on mortgages and less likely to downsize during unemployment spells. These results suggest that homeowners with wealthy parents find illiquidity of homeownership less problematic.

@@

\section{Transfers and the Rent-to-Own Transition}\label{app:rent_to_own}
A large literature documents that parental transfers increase the probability that renters become homeowners (e.g., \citealp{wold2024housing}; \citealp{Blickle2019}; \citealp{benetton2022dynastic}; \citealp{Guiso2002}; \citealp{Engelhardt1998}), I replicate this result to highlight the importance of transfers for housing outcomes.

I estimate
\begin{equation}	\label{eq:transferrent}
	\Pr\left(Own_{i,t+2}=1 \mid Own_{i,t}=0\right)
	= \alpha + \beta\,Transfer_{i,t+2}
	  + \gamma' X_{i,t} + \varepsilon_{i,t},
\end{equation}
where \(Own_{i,t}=0\) denotes a renter at time \(t\), \(Transfer_{i,t}=1\) if the household received a transfer between survey waves, and \(X_{i,t}\) includes log parental wealth, log household income and log household net worth, log household size, age, parental age, and dummies for  education, marital status, race, household composition changes, and state and year fixed effects. The sample is limited to renting households aged 25 to 44 who have never been observed as homeowners. Table \ref{tab:newowners} reports the estimated coefficients. These regressions are similar to those in \cite{Lee2018}, who also use the PSID.  It is important to note that transfer receipt is not exogenous---many transfers may be made precisely because the renter is buying a home.
